---
title: "FARSpckg -vignette"
author: "MW"
date: "November 13, 2017"
output: html_document
---

```{r, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# FARSpckg: Read, summarize and plot FARS data

This package lets the user read data from the US National Highway Traffic Safety Administration's Fatality Analysis Reporting System (FARS) saved as a CSV file. In addition several functions to summarize and plot FARS data are included. 

This vignette assumes you have downloaded the FARS data sets (as cvv.bz2) for the years 2013, 2014 and 2015 to your working directory from this [download link](https://d3c33hcgiwev3.cloudfront.net/_e1adac2a5f05192dc8780f3944feec13_fars_data.zip?Expires=1510704000&Signature=YwIRUwFBVVMh1FqVO5TnieqgFhA-tnqm-CYgO6QuW5LxUYeNzwE10yHrJ4-KIsZ0-Ja19TfbdKmwXy9IpAKsLS7PHYVuY5LKhAy1MwBj2~GrP-QiyX~-v89Lv9uQ3M3QEI6l3Ni4UKv9P1lHmW7RtBOWKBkio-qjbbZFwSzR6gg_&Key-Pair-Id=APKAJLTNE6QMUY6HBC5A). Otherwise you could also use the *fars_read* function to directly specify a filename.

## Workflow example

Start by loading the FARS data for 2013 and 2014.

```{r, echo=TRUE}
library(FARSpckg)
dat <- fars_read_years(2013:2014)
class(dat)
```

The result is a list with one entry for each requested year. Each entry looks something like this where is line represents one accident.
```{r, echo=TRUE}
dat[[1]]
```

The above function is used internally for the next function *fars_summarize_years* which summarizes the number of fatal injuries suffered in motor vehicle traffic crashes in each month for each year requested by the user.

```{r, echo=TRUE}
fars_summarize_years(2013:2014)
```

Finally, we have the function *fars_map_state* function which displays a basic map of where fatal injuries suffered in motor vehicle traffic crashes occured for a specific year in a specific state.

```{r, echo=TRUE}
fars_map_state(state.num=5, year=2013)
```
